% A workaround to allow relative paths in included subfiles
% that are to be compiled separately
% See https://tex.stackexchange.com/questions/153312/subfiles-inside-a-subfile-using-relative-paths
\providecommand{\main}{..}
\documentclass[\main/thesis.tex]{subfiles}

\begin{document}

\chapter{Discussion}

This project was focused on understanding how a large language model such as GPT2-XL changes onces it goes through
the process of ``specialization'' for a task such as knowledge graph generation. In order to understand the differences 
between the two models, we compared their performances for the GLUE benchmark set and arrived at a mixed conclusion. 

For the majority of the GLUE tasks, the models behave very similarly and have comparable scores. The only task where 
COMET seems to have struggled was with the Corpus of Linguistic Acceptability task. In this case, the un-specialized 
GPT2-XL seems to fare better by a significant margin. Upon analyzing the predictions for this task, we noticed that 
COMET seems to struggle with sentences in past tense. 

This points us in the right direction for conducting further research. Given the time constraints and the 
computational cost of training and testing LLMs, we had to limit this project to this GLUE task and even that took 
over one month of computation time on a power GPU to complete. 

In the future, it would be valuable to understand how newer models such as GPT-J and GPT-3 get change in case
they have been trained for knowledge graph generation. 
\end{document}